Data irregularities across five implicit learning articles by the same researchers

### 2021-08-19 docs/manuscript/manuscript.Rmd
update: merge brad's edit

---
title             : "Data irregularities across five implicit learning articles by the same researchers"
# title             : "Data irregularities across five implicit learning: Comments on @lola2012; @lola2020; @lola2021; @tzetzis2010; @tzetzis2015"
shorttitle        : "Data irregularities"

@@

## The articles in question

@@

In Articles 1-3, the groups differed with respect to the type of instruction received: implicit, explicit, or analogy. In Articles 4 and 5, a sequential group (see below for description) replaced the analogy group. All five experiments also included a control group that did not practice the task. Implicit instruction did not contain any explicit information for how to perform the task and the learners were asked to perform a distracting task like counting backwards while practicing to prevent them from acquiring declarative rules for performance. In contrast, explicit instruction consisted of direct verbal instructions for performing the task. Analogy instruction was considered a type of implicit instruction wherein an analogy or metaphor was provided to the learner. For example, "Imagine that the opponents' surface is covered with water. Send the ball where there is more water and no opponents at the court." [@lola2021, p. 9]. Sequential instruction involved receiving explicit instruction for the first half of training followed by implicit instruction for the second half of training. Across experiments, the authors predicted that implicit forms of instruction---implicit, analogy, and sequential---would be more effective than explicit instruction for motor and perceptual learning. This advantage was also predicted to be greater when testing was conducted in a high stress situation. In @lola2021 for instance, high stress was induced by falsely telling participants that the best performers would be selected for a draft to the national team. Further, it was often predicted that analogy or sequential instruction would offer improvements relative to implicit instruction.

The primary outcome measures used in these experiments were reaction time (Articles 1, 3, 4, and 5), response accuracy (Articles 1, 3, 4), and motor performance measured on a 4-point scale (Article 2). In addition, Articles 1 and 5 included a measure of state anxiety with the Competitive State Anxiety Inventory-2 (CSAI-2) [@tsorbatzoudis1998], and Article 2 had a measure of self-efficacy using a Likert scale. Articles 1, 3, 4, and 5 also analyzed the number of explicit rules recalled.

@@

None of the five articles in question included a link to a public repository where the data could be accessed. We first wrote (email sent February 11 2021) the corresponding author for Article 1 and asked if they would be willing to share the data for this experiment. The authors' response was that the data could not be shared as they were not finished with their analysis and were in the process of running different tests (A. Lola, personal communication, February 12 2021). We followed up this email (sent February 12 2021) by asking whether they would instead be willing to share the data from Articles 2 through 5 as these articles were less recent and presumably, all planned analyses had been completed. After a 2 week period with no response, we followed up on with a third email (sent February 26 2021) and reiterated our interest in obtaining their data from Articles 2 through 5. The authors' response was that they were unable to share data from any of these articles because in some cases they no longer had the data and in other cases they had plans to conduct further analyses (A. Lola, personal communication, March 2 2021).

Our first two requests did not include any indication that we were concerned about irregularities. Subsequently, in a fourth email (sent April 12 2021) we outlined our concerns for each article and once again reiterated our request to the authors to share any available data for any of the target articles. These requests were once again refused. The authors did address some specific concerns regarding Article 1 and provided some more general responses to our concerns; however, elements of their responses were inaccurate relative to the target articles. <!-- [MAYBE AN APPENDIX TO HIGHLIGHT SOME LIKE THE MENTION OF BONFERRONI YET NON USED BONFERONNI]. --> The authors did admit that some of the values reported in the target articles were incorrect but did not identify which values or articles. Despite this, the authors maintained that the data irregularities---identified in our email and described in this paper---do not impact the veracity of their analyses or conclusions (A. Lola, personal communication, April 22 2021). We illustrate below that the data *and* analyses reported in each of the articles reviewed are unreliable. Our extracted data and analysis scripts can be accessed at the following link: **[LINK]**.

## Effect size calculations and simulations

Means and standard deviations were extracted from each article for all measures and time points that were reported. Cohen's $d$ was calculated for each pairwise comparison using the \texttt{compute.es} package in \texttt{R}. To provide context, we simulated data with true effect sizes of $d$ = .80 and $d$ = 3.0 one million times each and report the range of effect sizes observed in those simulations. <!-- SHOULD WE NEED TO MENTION SAMPLE SIZE FOR SIMS AND WHY THAT VALUE WAS PICKED? -->

## Mathematically impossible data and granularity analysis

In two of the articles in question, it was clear that the some of the reported results were not mathematically possible based on the scale of measurement that was used. When outcomes were single item integers, such as the number of explicit rules recalled, we used a web application (<http://www.prepubmed.org/grimmer_sd/>) to conduct a granularity analysis called GRIMMER [@anaya2016]. Using GRIMMER it is possible to determine if specific means and standard deviations are possible for a given sample size. To be conservative, we specified that we did not know whether the standard deviation was calculated for the sample or population, nor whether ambiguous values were rounded up or down. Mean and standard deviation pairs that are mathematically possible are considered GRIMMER consistent, while mean and standard deviation pairs that are not mathematically possible are GRIMMER inconsistent. <!-- THIS IS THE SECTION I THINK WE MAY NEED TO ADD SOME EXPLANATIONS TO REGARDING THESE TESTS SO THE READER HAS AN UNDERSTANDING OF IT. I THINK WE COULD MODEL THIS OFF THE VAN DER ZEE ET AL PAPER AND THEIR GRANULARITY ERRORS SECTION -->

## Test statistics and $\eta^{2}$

Each of the articles reported only omnibus test statistics and then reported post-hoc analyses with symbols demarcating significant and non-significant differences. In response to our expressions of concern, the authors suggested that many of the issues were due to misprints in the articles. Specifically, they indicated that the reported means and standard deviations in their tables were incorrect and the root of the errors had to be from them outsourcing the formatting of their tables. The authors then insisted that despite these typographic errors, their discussion of the results and corresponding conclusions were still accurate (A. Lola, personal communication, April 22, 2021). However, the test statistics reported for many analyses were implausibly large and the authors often reported $\eta^{2}$ values associated with the omnibus test. Our examination of the reported $\eta^{2}$ values revealed that many were not only implausibly large, but also consistent with the reported pairwise effects.

@@

## Implausible effect sizes

@@

To evaluate the maximum range of plausible Cohen's $d$ statistics one might encounter from experiments similar to those reported in Articles 1-5, we conducted two simulations that each consisted of 1,000,000 experiments. For the first simulation we set the true effect size at $d$ = .80, the conventional benchmark for a "large" treatment effect. The largest effect size observed from the 1,000,000 simulated experiments was $d$ = 2.97 (see Figure \@ref(fig:fig1)). In the second simulation, we set the true effect size at $d$ = 3.0, an unrealistically large effect size that might rarely be encountered in the psychology literature. The maximum effect size observed in the 1,000,000 simulated experiments was $d$ = 6.6.

(ref:fig1-caption) Absolute Cohen's d estimates from Articles 1-4 plotted on a logarithmic scale. All pairwise comparisons have been included for all dependent measures in each experiment. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = .80 is illustrated by light blue and dark blue regions of the figure, reaching a maximum value of $d$ = 2.97. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = 3.0 is illustrated by the purple and dark blue regions of the figure, reaching a maximum value of $d$ = 6.86.

@@

<!-- PROBABLY NEED TO DO SOME MINOR TWEAKS TO FIG TO MAKE AXIS TEXT A BIT BIGGER. ALSO, NOT SURE WHAT COLOR CODE YOU PUT IN FOR THE BOTTOM SHADED AREA BUT I SEE YOU'VE DESCRIBED IT AS LIGHT BLUE IN THE CAPTION BUT IT LOOKS GREEN TO ME SO NOT SURE IF IT JUST ME AND COLORS OR NOT. -->

In the context of the maximum values observed in our simulations, all five articles in question reported implausibly large effect sizes. In Article 1, the smallest pre-test difference on the reaction measure was $d$ = 1.29, while the largest pre-test difference was $d$ = 35.32, although none of the groups were reported as significantly different in the article. The smallest post-intervention effect at any of the three time points was $d$ = 286.42, while the largest effect was $d$ = 3504.86.

A similar picture emerges when analyzing the accuracy data. All the pre-test differences were implausibly large (all $d$'s $\geq$ 2.52) despite being reported as not significantly different in the articles. Ten of the pairwise comparisons resulted in $d$'s $\geq$ 100 following treatment with the independent variables. The motor component data also revealed many post-treatment effect sizes that are implausibly large, ranging from $d$ = 1.16 to $d$ = 13.5. <!-- NOT SURE BUT IS IMPLAUSIBLY LARGE BEING THROWN AROUND TOO OFTEN THROUGHOUT THIS SECTION...SHOULD WE MAYBE JUST REPORT THE RANGE WITHOUT QUALIFYING IT LIKE THE PARAGRAPH BELOW? -->

In Article 2, post-intervention motor performance effect sizes ranged from $d$ = 3.1 to $d$ = 20.95. Similarly, post-intervention self-efficacy effect sizes ranged from $d$ = 1.79 to $d$ = 44.46. Similarly, in Article 3 post-intervention reaction time effect sizes ranged from $d$ = 2.28 to $d$ = 35.97. Continuing this pattern, post-intervention response accuracy effect sizes ranged from $d$ = 5.84 to $d$ = 29.7.

In Article 4, many response accuracy effect sizes were implausibly large beginning at pre-test, wherein effects ranged from $d$ = 2.53 to $d$ = 15.50. Nevertheless, all pre-test comparisons were reported as non-significant. Following intervention, the effect sizes ranged from $d$ = 23.13 to $d$ = 155.08.

The reaction time effect sizes were not implausibly large at any time point, ranging from $d$ = 0.0 to $d$ = .86. However, the authors reported an implausibly large effect size for the 4 (Group) x 3 (Time) ANOVA, $\eta^{2}$ = .94. Further, despite only one pairwise comparison being statistically significant, all post-intervention comparisons were reported as being significant in the article. <!-- IS THIS ALSO FOR ARTICLE 4, IF SO I THINK WE SHOULD COMBINE WITH PREVIOUS PARAGRAPH. I THINK IT BE GOOD TO HAVE A PARAGRAPH PER ARTICLE RATHER THAN A BUNCH OF SHORT PARAGRAPHS. THOUGHTS? -->

In Article 5 the authors did not report means and standard deviations for most of the analyses. However, $\eta^{2}$ effect sizes were reported and these ranged from $\eta^{2}$ = .52 to $\eta^{2}$ = .98. These effect sizes are discussed further below.

All the post-intervention effects reviewed above were directionally consistent with the researchers' expectations. The sometimes implausibly large pre-test effects were not expected, but also were not reported as significant.

## Impossible data and granularity analysis

@@

In Article 1, the mean and standard deviation pairs were GRIMMER inconsistent for all four groups. In Article 4, the mean and standard deviation pair was GRIMMER consistent for the explicit rules group (M = 4.8, SD = 1.78). The mean and standard deviation pairs for the remaining three groups were GRIMMER inconsistent. In Article 5, the mean and standard deviation pairs were GRIMMER consistent for three of the four groups if the standard deviations were calculated for the population rather than the sample. For two of the groups, however, they were consistent regardless which method of calculating the standard deviation was used. However, the results for the explicit group were GRIMMER inconsistent (M = 4.8, SD = 1.78).[^4]

[^4]: You may have noticed that the same mean and standard deviation pairing (M = 4.8, SD = 1.78) was classified as GRIMMER inconsistent for one paper and consistent for the other. This is because of sample size differences ($n$ = 20 for the consistent result and $n$ = 12 for the inconsistent one).

## Test statistics and $\eta^{2}$

Eta-squared ($\eta^{2}$) is calculated by dividing the sum of squares for the effect by the total sum of squares. It can be interpreted as analogous to $R^{2}$: It represents the total variation in the dependent measure that can be explained by a given main effect or interaction in an analysis of variance [@lakens2013]. Benchmarks have been suggested for small, medium, and large $\eta^{2}$ effect sizes as $\eta^{2}$ = .01, .06, and .14, respectively [@cohen1988]. Importantly, if a main effect of instruction type results in $\eta^{2}$ = .99, this suggests that 99% of the total variability in outcome measure can be explained by group assignment alone. Such a result is implausible.
<!-- IS THIS .99 JUST AN EXAMPLE? IF SO, MAYBE WE USE A MORE REALISTIC VALUE TO EXPLAIN THE INTERPRETATION. IF IT IS NOT AN EXAMPLE, THEN I THINK WE SHOULD REFER TO AN ARTICLE THAT HAS A .99 VALUE -->

The range of significant $F$-statistics reported across the five articles in question was $F$ = 28.60 to $F$ = 101,489.71. Only three $F$-statistics were under 100 (28.60, 30.73, and 91.75), and all three were for interactions.

@@

## Other oddities

Although the means and standard deviations for the explicit rules analysis were only reported in three of the articles in question, analyses were reported in Articles 1, 3, 4, and 5. The reported test statistic in these four articles was $F$ = 52.67, albeit with different degrees of freedom in Article 5 that reflected the different sample size in this experiment (48 versus 60 in Articles 1, 3, and 4). Articles 1-3 were published over a span of 6 years with reported samples sizes of 60 in Articles 1 and 3, and 80 in Article 2. Yet, the authors report identical means and standard deviations for the age of their participants in these three articles (see Table \@ref(tab:table1)). We assumed that each article was based on different samples as none of the articles mentioned using any previously published data.
<!-- I REMOVED THE F-STATISTIC ISSUE FROM ABOVE AS IT FELT A BIT REPETITIVE BEING IN TWO SECTIONS AND DECIDED TO LET IT LIVE JUST IN THIS ODDITIES SECTION INSTEAD. FEEL FREE TO SWITCH AROUND IF YOU THINK IT WOULD BE BEST HIGHER UP. -->

@@

Other irregularities, such a repeating $F$-statistic for all four analyses of explicit rules and the recurring age of participants, potentially reflect sloppiness more than expectation. Indeed, the authors have already admitted that some values reported in their tables were in error, and it seems errors occurred in each of the articles we have reviewed. These errors were pervasive and appear to have substantially affected the conclusions of the articles in question. At a minimum, the consistent reporting errors across these five articles seem to reflect excessive carelessness throughout the publication process. Even if the authors offer corrections, which they have suggested they intend to do[^5], many in the research community may find it difficult to trust any of these results.

>>>>

---
title             : "Data irregularities across five implicit learning: Comments on @lola2012; @lola2020; @lola2021; @tzetzis2010; @tzetzis2015"

shorttitle        : "Data irregularities"

@@

## The Articles in Question

@@

In Articles 1-3, the groups differed with respect to the type of instruction received: implicit, explicit, or analogy. In Articles 4 and 5, a sequential group (see below for description) replaced the analogy group. All five experiments also included a control group that did not practice the task. Implicit instruction did not contain any explicit information for how to perform the task and the learners were asked to perform a distracting task like counting backwards while practicing to prevent them from acquiring declarative rules for performance. In contrast, explicit instruction consisted of direct verbal instructions for performing the task. Analogy instruction was considered a type of implicit instruction wherein an analogy or metaphor was provided to the learner. For example, "Imagine that the opponents' surface is covered with water. Send the ball where there is more water and no opponents at the court." [@lola2021, p. 9]. Sequential instruction involved receiving explicit instruction for the first half of training followed by implicit instruction for the second half of training. Across experiments, the authors predicted that implicit forms of instruction---implicit, analogy, and sequential---would be more effective than explicit instruction for motor and perceptual learning. This advantage was also predicted to be greater when testing was conducted in a high stress situation. In @lola2021 for instance, high stress was induced by falsely telling participants that the best performers would be selected for a draft to the national team. Further, it was predicted that analogy or sequential instruction would offer improvements relative to implicit instruction.

The primary outcome measures used in these experiments were reaction time (Articles 1, 3, 4, and 5), response accuracy (Articles 1, 3, 4), and motor performance measured on a 4-point scale (Article 2). In addition, Articles 1 and 5 included a measure of state anxiety, the Competitive State Anxiety Inventory-2 (CSAI-2) [@tsorbatzoudis1998], and Article 2 had a measure of self-efficacy using a Likert scale. Articles 1, 3, 4, and 5 also analyzed the number of explicit rules recalled.

@@

None of the five articles in question included a link to a public repository where the data could be accessed. We first wrote (email sent February 10, 2021) the corresponding author for Article 1 and asked if they would be willing to share the data for this experiment. The authors' response was that the data could not be shared as they were not finished with their analysis and were in the process of running different tests (A. Lola, personal communication, February 12 2021). We followed up this email (sent February 12 2021) by asking whether they would instead be willing to share the data from Articles 2 through 5 as these articles were less recent and presumably, all planned analyses had been completed. After a 2 week period with no response, we followed up on with a third email (sent February 26 2021) and reiterated our interest in obtaining their data from Articles 2 through 5. The authors' response was that they were unable to share data from any of these articles because in some cases they no longer had the data and in other cases they had plans to conduct further analyses (A. Lola, personal communication, March 2 2021).

Our first two requests did not include any indication that we were concerned about irregularities. Subsequently, in a fourth email (sent April 12 2021) we outlined our concerns for each article and once again reiterated our request to the authors to share any available data for any of the target articles. These requests were once again refused. The authors did address some specific concerns regarding Article 1 and provided some more general responses to our concerns; however, elements of their responses were inaccurate relative to the target articles. The authors did admit that some of the values reported in the target articles were incorrect but did not identify which values or articles. Despite this, the authors maintained that the data irregularities---identified in our email and described in this paper---do not impact the veracity of their analyses or conclusions (A. Lola, personal communication, April 22 2021). We illustrate below that the data *and* analyses reported in each of the articles reviewed are unreliable. Our extracted data and analysis scripts can be accessed at the following link: **[LINK]**.

## Effect Size Calculations and Simulations

Means and standard deviations were extracted from each article for all measures and time points that were reported. Cohen's $d$ was calculated for each pairwise comparison using the \texttt{compute.es} package in \texttt{R}. To provide context, we simulated data from two groups of $n = 20$, consistent with groups sizes in Articles 1-4. We ran simulations with true effect sizes of $d = .80$ and $d = 3.0$ one million times each and report the range of effect sizes observed in those simulations.

## Mathematically Impossible Data and Granularity Analysis

In two of the articles in question, it was clear that the some of the reported results were not mathematically possible based on the scale of measurement that was used. When outcomes were single item integers (a granularity of 1), such as the number of explicit rules recalled, we used a web application (<http://www.prepubmed.org/grimmer_sd/>) to conduct a granularity analysis (GRIMMER) of reported means and standard deviations [@anaya2016]. GRIMMER builds off the original Granularity-Related Inconsistency of Means (GRIM) analysis [@brown2017], which leveraged the fact that the means of granular data are also granular. Given a data set of size $N$ and granularity $G$, only means of granularity $G/N$ are possible. Thus, all possible means for data of a given $G$ and $N$ can be enumerated, and only means that match these possibilities are considered GRIM-consistent. The GRIMMER analysis extends this test by also evaluating whether mean-standard deviation pairs are possible. First, the GRIM analysis is conducted to determine if the mean is GRIM-consistent. Next, lower and upper bounds of the standard deviation are calculated based on how many decimals ($D$) it is reported to by $(SD\pm^[0.5/10^{D}])^{2}$. Then all possible variances between these bounds are enumerated, converted back to standard deviations, and rounded to the nearest $D$ decimals. The reported standard deviation is checked for a match with any of these values. Finally, the mean-variance pair is compared to possible mean-variance pairings (the GRIMMER test handles sample sizes between 5 and 99).  Using GRIMMER it is possible to determine if specific mean and standard deviation pairs are possible for data of a given sample size. To be conservative, we specified that we did not know whether the standard deviation was calculated for the sample or population, nor whether ambiguous values were rounded up or down. Mean and standard deviation pairs that are mathematically possible are considered GRIMMER consistent, while mean and standard deviation pairs that are not mathematically possible are GRIMMER inconsistent.

## Eta-squared ($\eta^{2}$)

Each of the articles reported only omnibus test statistics and then reported post-hoc analyses with symbols demarcating significant and non-significant differences. In response to our expressions of concern, the authors suggested that many of the issues were due to misprints in the articles. Specifically, they indicated that the reported means and standard deviations in their tables were incorrect and the root of the errors had to be from them outsourcing the formatting of their tables. The authors then insisted that despite these typographic errors, their discussion of the results and corresponding conclusions were still accurate (A. Lola, personal communication, April 22, 2021). However, the test statistics reported for many analyses were implausibly large and the authors often reported $\eta^{2}$ values associated with the omnibus test. Our examination of the reported $\eta^{2}$ values revealed that, as with the reported pairwise comparisons, many were implausibly large.

@@

## Implausible Effect Sizes

@@

To evaluate the maximum range of plausible Cohen's $d$ statistics one might encounter from experiments similar to those reported in Articles 1-5, we conducted two simulations that each consisted of 1,000,000 experiments. For the first simulation we set the true effect size at $d$ = .80, the conventional benchmark for a "large" treatment effect. The largest effect size observed from the 1,000,000 simulated experiments was $d = 2.97$ (see Figure \@ref(fig:fig1)). In the second simulation, we set the true effect size at $d = 3.0$, an unrealistically large effect size that might rarely be encountered in the psychology literature. The maximum effect size observed in the 1,000,000 simulated experiments was $d = 6.6$.

(ref:fig1-caption) Absolute Cohen's d estimates from Articles 1-4 plotted on a logarithmic scale. All pairwise comparisons have been included for all dependent measures in each experiment. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d = .80$ is illustrated by light green and dark blue regions of the figure, reaching a maximum value of $d = 2.97$. The range of observed values from a simulation of 1,000,000 experiments with a true effect of $d$ = 3.0 is illustrated by the purple and dark blue regions of the figure, reaching a maximum value of $d = 6.6$.

@@

In the context of the maximum values observed in our simulations, all five articles in question reported implausibly large effect sizes. In Article 1, the smallest pre-test difference on the reaction measure was $d = 1.29$, while the largest pre-test difference was $d = 35.32$, although none of the groups were reported as significantly different in the article. The smallest post-intervention effect at any of the three time points was $d = 286.42$, while the largest effect was $d = 3504.86$. A similar picture emerges when analyzing the accuracy data. All the pre-test differences were improbably large (all $d$'s $\geq$ 2.52) despite being reported as not significantly different in the articles. Ten of the pairwise comparisons resulted in $d$'s $\geq$ 100 following treatment with the independent variables. The motor component data revealed post-treatment effect sizes ranging from $d = 1.16$ to $d = 13.5$.

In Article 2, post-intervention motor performance effect sizes ranged from $d = 3.1$ to $d = 20.95$. Similarly, post-intervention self-efficacy effect sizes ranged from $d = 1.79$ to $d = 44.46$. Likewise, in Article 3 post-intervention reaction time effect sizes ranged from $d = 2.28$ to $d = 35.97$. Continuing this pattern, post-intervention response accuracy effect sizes ranged from $d = 5.84$ to $d = 29.7$.

In Article 4, many response accuracy effect sizes were implausibly large beginning at pre-test, wherein effects ranged from $d = 2.53$ to $d = 15.50$. Nevertheless, all pre-test comparisons were reported as non-significant. Following intervention, the effect sizes ranged from $d = 23.13$ to $d = 155.08$. The reaction time effect sizes were not implausibly large at any time point, ranging from $d = 0.0$ to $d = .86$. However, the authors reported an implausibly large effect size for the 4 (Group) x 3 (Time) ANOVA, $\eta^{2}$ = .94. Further, despite only one pairwise comparison being statistically significant, all post-intervention comparisons were reported as being significant in the article.

In Article 5 the authors did not report means and standard deviations for most of the analyses. However, $\eta^{2}$ effect sizes were reported and these ranged from $\eta^{2}$ = .52 to $\eta^{2}$ = .98. These effect sizes are discussed further below. All the post-intervention effects reviewed above were directionally consistent with the researchers' expectations. The sometimes implausibly large pre-test effects were not expected, but also were not reported as significant.

## Impossible Data and Granularity Analysis

@@

In Article 1, the mean and standard deviation pairs were GRIMMER inconsistent for all four groups. In Article 4, the mean and standard deviation pair was GRIMMER consistent for the explicit rules group ($M = 4.8$, $SD = 1.78$). The mean and standard deviation pairs for the remaining three groups were GRIMMER inconsistent. In Article 5, the mean and standard deviation pairs were GRIMMER consistent for three of the four groups if the standard deviations were calculated for the population rather than the sample. For two of the groups, however, they were consistent regardless which method of calculating the standard deviation was used. However, the results for the explicit group were GRIMMER inconsistent ($M = 4.8$, $SD = 1.78$).[^4]

[^4]: You may have noticed that the same mean and standard deviation pairing ($M = 4.8$, $SD = 1.78$) was classified as GRIMMER inconsistent for one paper and consistent for the other. This is because of sample size differences ($n = 20$ for the consistent result and $n = 12$ for the inconsistent one).

## Eta-squared ($\eta^{2}$)

Eta-squared ($\eta^{2}$) is calculated by dividing the sum of squares for the effect by the total sum of squares. It can be interpreted as analogous to $R^{2}$: It represents the total variation in the dependent measure that can be explained by a given main effect or interaction in an analysis of variance [@lakens2013]. Benchmarks have been suggested for small, medium, and large $\eta^{2}$ effect sizes as $\eta^{2}$ = .01, .06, and .14, respectively [@cohen1988]. Importantly, if a main effect of instruction-type results in $\eta^{2}$ = .99, as we will see commonly below, this suggests that $99\%$ of the total variability in the outcome measure can be explained by group assignment alone. Such a result is implausible.

@@

## Other Oddities

Although the means and standard deviations for the explicit rules analysis were only reported in three of the articles in question, analyses were reported in Articles 1, 3, 4, and 5. The reported test statistic in these four articles was $F = 52.67$, albeit with different degrees of freedom in Article 5 that reflected the different sample size in this experiment (48 versus 60 in Articles 1, 3, and 4). Articles 1-3 were published over a span of 6 years with reported samples sizes of 60 in Articles 1 and 3, and 80 in Article 2. Yet, the authors report identical means and standard deviations for the age of their participants in these three articles (see Table \@ref(tab:table1)). We assumed that each article was based on different samples as none of the articles mentioned using any previously published data.

@@

Other irregularities, such as a repeating $F$-statistic for all four analyses of explicit rules and the recurring age of participants, potentially reflect sloppiness more than expectation. Indeed, the authors have already admitted that some values reported in their tables were in error, and it seems errors occurred in each of the articles we have reviewed. These errors were pervasive and appear to have substantially affected the conclusions of the articles in question. At a minimum, the consistent reporting errors across these five articles seem to reflect excessive carelessness throughout the publication process. Even if the authors offer corrections, which they have suggested they intend to do[^5], many in the research community may find it difficult to trust any of these results.
